Sorting algorithms: how are they affected by memory faults?

### 2019-12-15 paper/MaterialsAndMethods.tex

From that, we develop a Python script to generate custom inputs. The sequence has two characteristics: (\textit{i}) the sequence size, and (\textit{ii}) the probability of failure. For this work, the generated input files are shown in the Table \ref{table-input-data} below.

\begin{table}[htbp]
    \caption{Generated input data.}
    \begin{center}
    \begin{tabular}{|c|c|c|c|}
    \hline
    \multirow{1}{*}{\textbf{Input}}&\textbf{Sequence Size}&\textbf{Probability of Failure} \\
    \hline
    Input A & 100 & 1\% \\
    Input B & 100 & 2\% \\
    Input C & 100 & 5\% \\
    \hline
    Input D & 1000 & 1\% \\
    Input E & 1000 & 2\% \\
    Input F & 1000 & 5\% \\
    \hline
    Input G & 10000 & 1\% \\
    Input H & 10000 & 2\% \\
    Input I & 10000 & 5\% \\
    \hline
    \end{tabular}
    \label{table-input-data}
    \end{center}
\end{table}

@@

Figure \ref{input-file-example} shows an example of an input file generated by the Python script. The first number of the sequence (\texttt{0.01}) is the probability of memory failure when sorting. The second number (\texttt{100}) means the size of the integers sequence used by sorting. The rest of the numbers indicates the sequence itself.

>>>>

@@

The input data shown in the Figure \ref{input-file-example} is divided as follows:

\begin{itemize}
    \item \textit{Probability of Failure}: the first number of the sequence (\texttt{0.01}) is the probability of memory failure when sorting;
    \item \textit{Sequence size}: the second number (\texttt{100}) means the size of the integers sequence used by sorting;
    \item \textit{Sequence}: the rest of the numbers indicates the sequence itself.
\end{itemize}

We run 1000 times each algorithm with each file listed in Table \ref{table-input-data} below. We chose this number of executions to reduce possible noise in the generated data. Table \ref{table-input-data} shows 9 different input files with its sequence sizes and probabilities of failure. Plus, the number of executions and each sorting algorithm.

For example, Input A showed in that table with its characteristics was executed 1000 times for each algorithm -  bubblesort, quicksort, mergesort, and insertion sort - totalizing 4000 executions.

\begin{table}[H]
    \caption{Generated input data.}
    \begin{center}
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    \textbf{Input} & \textbf{Sequence Size} & \textbf{Probability of Failure} & \textbf{Executions} & \textbf{Algorithm} \\ 
    \hline
    \multirow{4}{*}{Input A} & \multirow{4}{*}{100} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input B} & \multirow{4}{*}{100} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input C} & \multirow{4}{*}{100} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input D} & \multirow{4}{*}{1000} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input E} & \multirow{4}{*}{1000} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input F} & \multirow{4}{*}{1000} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input G} & \multirow{4}{*}{10000} & \multirow{4}{*}{1\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input H} & \multirow{4}{*}{10000} & \multirow{4}{*}{2\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \multirow{4}{*}{Input I} & \multirow{4}{*}{10000} & \multirow{4}{*}{5\%} & \multirow{4}{*}{1000} & Bubblesort \\ 
    & & & & Quicksort \\
    & & & & Mergesort \\
    & & & & Insertion Sort \\
    \hline
    \end{tabular}
    \label{table-input-data}
    \end{center}
\end{table}

Each execution generates an output file. Then, we generate 4000 outputs for each combination of sequence size, and the probability of failure, totalizing 36000 files. Figure \ref{output-file-example} shows an output file.

\begin{verbbox}[\mbox{}]
[1]  9 48 37 6 26 7 24 44 17 50 48 30 49 33 22 13 42 29 39 13 19 13 9 28 34 1
33 27 14 45 48 40 11 17 6 50 9 44 20 16 37 45 23 14 38 29 10 49 44 46 35 45  
15 2 22 1 46 40 8 48 23 23 32 35 3 15 8 36 17 24 27 48 28 5 28 50 44 4 25 6 9
1 11 44 26 50 44 12 7 20 30 20 37 20 6 8 13 15 20 49
[2]  1 1 1 2 3 4 5 6 6 6 6 7 7 8 8 8 9 9 9 9 10 11 11 12 13 13 13 13 14 14 15 
15 15 16 17 17 17 19 20 20 20 20 22 22 23 23 23 24 24 25 26 26 27 27 28 28 28 
29 29 30 30 32 33 33 34 35 35 36 37 37 37 38 39 40 40 42 44 44 44 44 44 44 20 
45 45 45 46 46 48 48 48 48 48 49 49 50 49 50 50 50
[3]  1 1 1 2 3 4 5 6 6 6 6 7 7 8 8 8 9 9 9 9 10 11 11 12 13 13 13 13 14 14 15 
15 15 16 17 17 17 19 20 20 20 20 20 22 22 23 23 23 24 24 25 26 26 27 27 28 28 
28 29 29 30 30 32 33 33 34 35 35 36 37 37 37 38 39 40 40 42 44 44 44 44 44 44 
45 45 45 46 46 48 48 48 48 48 49 49 49 50 50 50 50
[4]  82
\end{verbbox}

\begin{figure}[H]
    \centering
    \fbox{
    \theverbbox
    }
    \caption{Example of output file.}
    \label{output-file-example}
\end{figure}

The output file gives four essential data, as enumerated below:
\begin{itemize}
    \item \textit{[1]}: the original sequence of integers contained in the input file;
    \item \textit{[2]}: the sequence processed by the sorting algorithm under the memory fault model;
    \item \textit{[3]}: the sequence sorted correctlty;
    \item \textit{[4]}: the size of the largest sorted subsequence in [2]. This number can be interpreted as the quality of sorting. As higher, most successful was the sorting operation.
\end{itemize}

We develop Python scripts to generate input files and analyze our data. To be continued...
